\documentclass[11pt]{article}

% ---- theme variables (passed via pandoc -V) ----
\newcommand{\ThemeName}{AI / ML — Index}
\newcommand{\ThemeKicker}{AI/ML Track}
\newcommand{\ThemeNumber}{}

\usepackage{interviewstyle}
\setthemecolor{7c3aed}{a78bfa}

% pandoc-required plumbing
\providecommand{\tightlist}{\setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}}
\setlength{\emergencystretch}{3em}
\providecommand{\pandocbounded}[1]{#1}

% syntax-highlighting macros emitted by pandoc (skylighting)


\begin{document}

\makecoverpage

\thispagestyle{empty}
{\hypersetup{hidelinks}\tableofcontents}
\clearpage

\section{AI / ML --- Interview Notes}\label{ai--ml--interview-notes}

From classical machine learning to deep learning, Transformers, and building real products with LLMs --- first-principles notes in the same structure as the core topics, with hand-built diagrams and magazine PDFs.

\subsection{Files}\label{files}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.0733}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5683}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1054}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2530}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\#
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Topic
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Notes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Magazine PDF
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & Machine Learning Fundamentals & \href{01-ml-fundamentals.md}{md} & \href{ML-Fundamentals-Magazine.pdf}{PDF} \\
2 & Classic ML Algorithms & \href{02-classic-ml-algorithms.md}{md} & \href{Classic-ML-Algorithms-Magazine.pdf}{PDF} \\
3 & Deep Learning \& Neural Networks & \href{03-deep-learning.md}{md} & \href{Deep-Learning-Magazine.pdf}{PDF} \\
4 & Transformers \& LLMs & \href{04-transformers-and-llms.md}{md} & \href{Transformers-and-LLMs-Magazine.pdf}{PDF} \\
5 & LLM Applications \& GenAI & \href{05-llm-applications-genai.md}{md} & \href{LLM-Applications-Magazine.pdf}{PDF} \\
6 & MLOps \& ML System Design & \href{06-mlops-and-ml-system-design.md}{md} & \href{MLOps-ML-System-Design-Magazine.pdf}{PDF} \\
\end{longtable}

= the architecture behind modern AI --- highest-signal for AI/ML interviews.

\subsection{What each covers}\label{what-each-covers}

\begin{itemize}
\tightlist
\item
  \textbf{01 Fundamentals} --- learning paradigms, bias--variance, regularization, evaluation metrics, gradient descent.
\item
  \textbf{02 Classic Algorithms} --- regression, trees, random forests, gradient boosting (XGBoost/LightGBM), SVM, k-NN, Naive Bayes, k-Means, PCA, ensembles.
\item
  \textbf{03 Deep Learning} --- neurons, backprop, activations, optimizers, CNNs, RNNs/LSTMs, batch norm, dropout, embeddings.
\item
  \textbf{04 Transformers \& LLMs} --- self-attention (Q/K/V), the Transformer, tokenization, BERT vs GPT, pretraining, RLHF/DPO, LoRA, quantization, MoE, decoding, hallucination.
\item
  \textbf{05 LLM Applications} --- prompt engineering, RAG (chunking, embeddings, vector DBs, reranking), agents (tools, function calling, ReAct), evaluation, guardrails, cost/latency.
\item
  \textbf{06 MLOps \& ML System Design} --- feature stores, distributed training, serving, drift, the ML-design framework, and worked designs (recommendation, CTR, feed ranking).
\end{itemize}

\subsection{Reading order}\label{reading-order}

\texttt{01\ →\ 02\ →\ 03\ →\ 04\ →\ 05}, with \texttt{06} as the systems capstone. For AI-focused roles, prioritize \textbf{04 + 05 + 06}.

\subsection{\texorpdfstring{Each \texttt{.md} has a coherent analogy world}{Each .md has a coherent analogy world}}\label{each-md-has-a-coherent-analogy-world}

School (fundamentals) \(\cdot\) a panel of doctors (algorithms) \(\cdot\) a learning factory line (deep learning) \(\cdot\) a newsroom of editors (Transformers) \(\cdot\) a research assistant in a library (LLM apps) \(\cdot\) an airport ops center (MLOps).

\begin{quote}
Preview \& edit any edition live: \texttt{cd\ ../.magazine\ \&\&\ python3\ serve.py} \(\rightarrow\) http://localhost:8000
\end{quote}

\end{document}
